---
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Known feature/bug of RNotebooks - Notebooks are executed in the same directory as the file but we want to set the working directory to the root directory.

```{r setup}
# Correct RNotebook working directory bug 
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/cfull/OneDrive/Desktop/suq21-5210-w2-pa2"))
```

<div style="float: right; width: 90%;">

```{r}
# Course: 5210 Communicating Data
# Purpose: Participation Assignment #4
# Date: July 22, 2021
# Author: Chequala Fuller

```


# Load Libraries

```{r}
# Load libraries

# For wrangling, subseting, and ploting
library(tidyverse)

# For correlation test of multiple variables
library(Hmisc)       

# Tidy cross tabs
library(janitor)       

# To calculate multinomial confidence intervals for factor variables
library(MultinomialCI)  

# To format html tables
library(htmlTable)     
