---
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

Known feature/bug of RNotebooks - Notebooks are executed in the same directory as the file but we want to set the working directory to the root directory.

```{r setup}
# Correct RNotebook working directory bug 
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/cfull/OneDrive/Desktop/suq21-5210-w3-quick-project-1"))
```

<div style="float: right; width: 90%;">

```{r}
# Course: 5210 Communicating Data
# Purpose: Participation Assignment #3
# Date: July 17, 2021
```

# Load Packages

```{r, echo = TRUE, warning = FALSE, error = FALSE}
# Load packages

# The tidyverse package contains ggplot2, dplyr, and several other packages we will use
library(tidyverse)

# The gridExtra package contains grid.arrange function used to combine plots in the same window
library(gridExtra)

# The janitor package contains tidyverse functions for cross-tables
library(janitor)

# The knitr package contains some table formating functions
library(knitr)

# The GGally package contains a custom correlation plot we will use
library(GGally)

# The scales package contains scale_y_continues for plot label formatting
library(scales)

# ggannotate
library(ggannotate)
```

# Load Data

```{r}
king_sales <- read_csv("data/qp1_data.csv")
```

# Base EDA Step 1 - Uni-variate Non-graphical Data

```{r}
# View the data
head(king_sales, 10)
```
**Comments**

* Data appears tidy:
  + Each column is a variable and the data is the same type
  + There are 13 variables
  + There are no duplicative variables
  + Each row is an observation for the sale of a specific home


## Data Structure

```{r}
# convert bedrooms, waterfront and view variables from numerical to factor
names <- c('waterfront', 'view', 'condition')
king_sales[,names] <- lapply(king_sales[,names], factor)

# Check the structure of the data
str(king_sales)
```
**Comments**

* All variables were initially numerical
  + waterfront, view, grade, and condition were converted to factor
* Each variable is unordered

## Descriptive statistics

```{r}
# Summary statistics of data
summary(king_sales)
```
**Comments**

* Data appears balanced across number of bedrooms

* Price, sqft_living, and sqft_lot variables appear to be right skewed (asymmetrical)

  + mean != median and mean >> median
  + in these cases, we'll use the median
  + yr_renovated is 84.4 for mean, 2015 for max but 0 for min, 1st Qu., median, and 3rd Qu. (that is alerting) - what could be going on here?
  
* Bedrooms, bathrooms and floors variables all appear to be symmetrical

  + mean = median
  
* For price, is 7700000 an outlier?

  + significantly higher than 3rd Qu. 645000

* For bedrooms, is 33 an outlier?

  + significantly higher than 3rd Qu. 4
  
* For bathrooms, is 8 an outlier?

  + higher than 3rd Qu. 2.5
  
* For sqft_living, is 13540 an outlier?

  + significantly higher than 3rd Qu. 2550
  
* For sqft_lot, is 1651359 an outlier?

  + significantly higher than 3rd Qu. 10688
  
**Questions**

* yr_renovated is 84.4 for mean, 2015 for max but 0 for min, 1st Qu., median, and 3rd Qu. What could be going on here?

* Is there really a specific home with 33 bedrooms and 8 bathrooms or could this be an anomaly?  


# Base EDA Step 2: Uni-variate Graphical

* In this step, we will examine each variable individually

## Categorical/Factor Variables:

* Take a look at all categorical variables together

```{r}

# Only use bar graphs
grid.arrange(
  
# waterfront
ggplot(data = king_sales, mapping = aes(x = waterfront)) +
  geom_bar(),
  
# view
ggplot(data = king_sales, mapping = aes(x = view)) +
  geom_bar(),

# condition
ggplot(data = king_sales, mapping = aes(x = condition)) +
  geom_bar(),
  

ncol = 1)
```

**Comments**

* Most homes sold are not waterfront locations and have little to no view. On the other hand the conditions when sold were mostly 3 and 4 (mediocre) for King County.


## Quantitative Variables

### Price

```{r}
# Use grid.arrange() to display multiple graphs together
grid.arrange(

# Create a histogram of price
ggplot(data = king_sales, mapping = aes(x = price)) +
  geom_histogram(color = "violetred", alpha = 0.4, adjust = 2),

# Add a boxplot
ggplot(data = king_sales, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = price), color = "violetred", alpha = 0.5) + 
  coord_flip(), # flip to make the x-axis the same

# Set the number of columns
ncol = 1)
```


**Comments**

* Confirm the price is still slightly right skewed as suspected

  + There were a number of outliers, which heavily skewed the mean.
  
  + Median is best to use for the analysis of this variable
  
* Most homes sold lie within the $300,000 to $600,000 price range


### Bedrooms

```{r}
# Use grid.arrange() to display multiple graphs together
grid.arrange(

# Create a histogram of price
ggplot(data = king_sales, mapping = aes(x = bedrooms)) +
  geom_histogram(binwidth = 1, color = "maroon4", alpha = 0.4, adjust = 2),

# Add a boxplot
ggplot(data = king_sales, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = bedrooms), color = "maroon4") + 
  coord_flip(), # flip to make the x-axis the same

# Set the number of columns
ncol = 1)
```

**Comments**

* Confirm that after eliminating outliers, the bedrooms variable is more normally distributed

  + Okay to use the mean

* The majority of homes sold in King Country were commonly 3 and 4 bedroom homes.

**Questins**

* Are the number of bedrooms and sqft_living correlated?

* Are the bedrooms and sqft_living relative to price?


### Bathrooms

```{r}
# Use grid.arrange() to display multiple graphs together
grid.arrange(

# Create a histogram of price
ggplot(data = king_sales, mapping = aes(x = bathrooms)) +
  geom_histogram(binwidth = 1, color = "plum", alpha = 0.4, adjust = 2),

# Add a boxplot
ggplot(data = king_sales, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = bathrooms), color = "plum") + 
  coord_flip(), # flip to make the x-axis the same

# Set the number of columns
ncol = 1)
```

**Comments**

* Even after eliminating outliers, bathrooms variable is still slightly right skewed but not a severe

  + Best to use the median

* Majority of homes sold had 2 bathrooms.


### Square Footage (Living)

```{r}
# Use grid.arrange() to display multiple graphs together
grid.arrange(

# Create a histogram of sqft_living
ggplot(data = king_sales, mapping = aes(x = sqft_living)) +
  geom_histogram(color = "darkorchid", alpha = 0.4, adjust = 2),
  

# Add a boxplot
ggplot(data = king_sales, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = sqft_living), color = "darkorchid") + 
  coord_flip(), # flip to make the x-axis the same

# Set the number of columns
ncol = 1)
```

**Comments**

* Confirm that sqft_living is normally distributed with a slight right skew

  + Okay to use the mean

* Majority of homes sold are within 1500 sqft to 2500 sqft, with multiple outliers

  + The outliers are skewing the mean so median is best to use for the analysis of this variable
  
  
### Square Footage (Lot)
```{r}
# Use grid.arrange() to display multiple graphs together
grid.arrange(

# Create a histogram of sqft_lot
ggplot(data = king_sales, mapping = aes(x = sqft_lot)) +
  geom_histogram(color = "magenta4", alpha = 0.4, adjust = 2),
  

# Add a boxplot
ggplot(data = king_sales, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = sqft_lot), color = "magenta4") +
  coord_flip(), # flip to make the x-axis the same

# Set the number of columns
ncol = 1)
```

**Comments**

* Confirms that sqft_lot is asymmetrical and right skew as suspected
  
  + Since data is skewed, we will use the median instead of mean

**Question**

* Does the sqft_lot have less of an effect on price than sqft_living?


### Year Built
```{r}
# Use grid.arrange() to display multiple graphs together
grid.arrange(

# Create a histogram of yr_built
ggplot(data = king_sales, mapping = aes(x = yr_built)) +
  geom_histogram(color = "mediumpurple3", alpha = 0.4, adjust = 2),
  

# Add a boxplot
ggplot(data = king_sales, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = yr_built), color = "mediumpurple3") +
  coord_flip(), # flip to make the x-axis the same

# Set the number of columns
ncol = 1)
```
**Comments**

* Confirm yr_built is assymetrical and left skewed

  + Did not eliminate year values. Especially after 2000, which create the outlier effect on the mean
  
  + Use the median
  
**Questions**

* Would eliminating years > 2000 improve or worsen the model?


# Base EDA Step 3: Multi-variate Non-graphical

## Categorical

```{r}
# Contingnecy/Cross tabulation for multiple variables

king_sales %>% 
  tabyl(bedrooms, grade) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  kable()
```
```{r}
# Proportion contingency/cross tabulation for bedrooms and waterfront

king_sales %>% 
  tabyl(bedrooms, grade) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # create proportions
  adorn_rounding(2) %>%  # round decimal two places
  kable()
```

```{r}
# Proportional contingency/cross tabulation for bedrooms and sqft_living
king_sales %>% 
  tabyl(bedrooms, waterfront) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all") %>% # create proportions
  adorn_rounding(4) %>%  # round decimal two places
  kable()
```

**Comments**

* Confirms the number of bedrooms on waterfront or view are not consistent across

  + Most sales fall under those that are not located on waterfront
  + We'll need to model these
  + We shouldn't analyze houses with 0, 7, 8, 9, 10, 11, and 33 bedrooms. Those have low counts

**Questions**

* Need to verify price as it relates to number of bathrooms and sqft_living


## Quantitative

```{r}
# Correlation table

king_sales %>% 
  select(is.numeric) %>% 
  cor() %>% 
  round(2) %>% 
  kable(align = 'c', caption = "Numeric Variables Correlation Table")
```

**Comments**

* As suspected, there is no significance in the sqft_lot and yr_renovated variables

  + We will eliminate the variables with little to no correlation at all from the analysis
  + Eliminate sqft_lot and yr_renovated
  
* There are large positive correlations between sqft_living/bathrooms (0.69), sqft_living/grade (0.67), sqft_living/bedrooms (0.61), bathrooms/grade (0.60), and price/grade (0.59) 

**Questions**

* Does positive correlation indicate causation?

* Does price increase with sqft_living and grade?


# Base EDA Step 4: Multi-variate Graphical

## Step 4.1: Categorical

* Distribution or summary of quantitative and categorical variable, bar graphs, and multiple boxplots

```{r}
# Use grid.arrange() so we can see all uni-variate bar graphs together

grid.arrange(
  
  king_sales %>% 
    ggplot(mapping = aes(x = sqft_living, fill = waterfront)) +
    geom_bar(position = "dodge") +
    coord_flip(),
  
  king_sales %>% 
    ggplot(mapping = aes(x = sqft_living, fill = view)) + 
    geom_bar(position = "dodge") +
    coord_flip(),
  
  king_sales %>% 
    ggplot(mapping = aes(x = sqft_living, fill = condition)) + 
    geom_bar(position = "dodge") +
    coord_flip(),
  
  ncol = 2
)
```

**Comments**

* Homes with no waterfront location nor view are more prominent

**Questions**

* Have more homes sold without waterfront or views in King County because of price?



## Step 4.1: Categorical

```{r}
# More detail on bedrooms and view

grid.arrange(
  # Cluster of counts
  king_sales %>% 
    ggplot(mapping = aes(x = bedrooms, fill = view)) + 
    geom_bar(position = "dodge") +
    coord_flip(),
  
  # Proportion of counts
  king_sales %>% 
    ggplot(mapping = aes(x = bedrooms, fill = view)) +
    geom_bar(position = "fill") +
    coord_flip(),
  
ncol = 1)
```


## Step 4.2: Quantitative

```{r}
# Scatterplots using grid.arrange() to see all quantitave variables of interest together

king_sales %>% 
  ggplot(mapping = aes(x = sqft_living, y = price)) +
  geom_point()
```

**Comments**

* This is pattern is quite clustered from 290 sqft to 5000 sqft. Then it disperses.


## Step 4.3: Categorical and Quantitative

```{r}
# Use ggpairs from the GGally package
king_sales %>% 
  select(-c(sqft_lot, yr_renovated)) %>% 
  ggpairs(columns = c(1:4, 6:8))
```

**Questions**

* Look closer at the following variable relationships:

  + sqft_living and bathrooms
  + sqft_living and grade
  + sqft_living and bedrooms
  + bathrooms and grade
  + price and grade

```{r}
# follow up on correlation graphs

grid.arrange(
  
  # sqft_living and bathrooms
  king_sales %>% 
    ggplot(mapping = aes(x = sqft_living, y = bathrooms)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE),
  
  # sqft_living and bedrooms
  king_sales %>% 
    ggplot(mapping = aes(x = sqft_living, y = bedrooms)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE),
  
  ncol = 1)
```

```{r}

grid.arrange(
  
  # sqft_living and grade
  king_sales %>% 
    ggplot(mapping = aes(x = sqft_living, y = grade)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE),
  
  # bathrooms and grade
  king_sales %>% 
    ggplot(mapping = aes(x = bathrooms, y = grade)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE),
  
  
  ncol = 1)
```

```{r}
grid.arrange(
  
  # price and grade
  king_sales %>% 
    ggplot(mapping = aes(x = bathrooms, y = grade)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE),
  

ncol = 1)
```

**Comments**

* There is a similar pattern between price and sqft_living across both grade and number of bathrooms 

  + It appears the price of sale increases as sqft_living and grade quality increase
  
  + Similarly, price also increases as sqft_living and bathrooms increase


# Detailed EDA - Questions Raised in Base EDA

## Verify price as it relates to the number of bathrooms. Does it differ?


```{r}
# facet histogram - price and grade
king_sales %>% 
  ggplot(mapping = aes(x = price)) +
  geom_histogram() +
  facet_wrap( ~ grade)
```

**Comments**

* The pattern holds
  

## Does the price differ across various grade?

```{r}

# facet histograms - sqft_living and bedrooms
king_sales %>% 
  ggplot(mapping = aes(x = sqft_living)) +
  geom_histogram() +
  facet_wrap( ~ bedrooms)
```

**Comments**

* The pattern holds. Let's explore this relationship further.


## Does price differ across sqft_living and grade? 

```{r}
# facet histograms - price
king_sales %>% 
  ggplot(mapping = aes(x = sqft_living, y = price)) +
  geom_jitter(width = 0.5) +
  facet_wrap( ~ grade)

king_sales %>% 
  ggplot(mapping = aes(x = sqft_living, y = price)) +
  geom_jitter(width = 0.5) +
  geom_smooth(method = "lm", se = FALSE)
```

# Summary

* Most houses sold in King County do not come with waterfront locations or good views at that.

* The King County housing market appears to be influenced by features including the number of square footage of living space, the grade quality of a home, as well as the number of bathrooms included.

* Pricing appears to be consistent across sqft_living and grade.


# Create Professional Visuals

```{r}
# Plot price and year
price_yr_built <- king_sales %>% 
  ggplot(mapping = aes(x = yr_built, y = price, color = yr_built)) +
  geom_col(postion = "fill") +
  labs(x = "Year Built",
       y = "Price of Homes Built",
       title = "The Number of Homes Built from 1900 - 2000",
       subtitle = "Common number of bedrooms per home depending on the era") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))


# View price_yr_built plot
price_yr_built
```

```{r}
# Bar graph for number of bedrooms and price of homes sold
price_grade_plot <- king_sales %>% 
  ggplot(mapping = aes(x = grade, y = price, fill = waterfront)) + 
  geom_col(position = "dodge")  + 
  labs(x = "Grade Quality",
       y = "Sales Price", 
       title = "King County Housing Prices based on Quality of Home",
       subtitle = "Is a waterfront view a determining factor?") + 
  scale_y_continuous(labels = dollar) + # add dollar sign and decimal to y-axis values
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

# View plot
price_grade_plot
```

# PA4


```{r}
# recreate the price and grade plot from qp1 data
rec_price_grade_plot <- king_sales %>% 
  ggplot(mapping = aes(x = grade, y = price)) + 
  geom_line(stat = "identity")  + 
  geom_point(color = "blue3") +
  scale_x_continuous(breaks = seq(0, 13, 2), lim = c(0, 13)) +
  labs(x = "grade",
       y = "price", 
       title = "Increasing Price with Higher Grade")

# View plot
rec_price_grade_plot
```

```{r}
# redesign it

red_price_grade_plot <- king_sales %>% 
  ggplot(mapping = aes(x = grade, y = price)) + 
  geom_line(stat = "identity")  + 
  geom_point(color = "") +
  scale_x_continuous(breaks = seq(0, 13, 2), lim = c(0, 13)) +
  scale_y_continuous(labels = dollar) + # add dollar sign and decimal to y-axis values
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  labs(x = "Grade",
       y = "Sales Price", 
       title = "Higher Grade Yields Increasing Sales Prices")

# View plot
red_price_grade_plot
```


# Save Visuals

```{r}
# Save previous graphs to use as professional visuals in the memo
ggsave(filename = "suq21-5210-w4-pa3-price_yr_built.png", plot = price_yr_built)
ggsave(filename = "suq21-5210-w4-pa3-price_grade_plot.png", plot = price_grade_plot)
```


  


